---
title: "Assignment 3 ggplot2"
author:
- familyname: Li
  othernames: Siyi
  address: Monash Universty
  email: "slii0048@student.monash.edu"
  correspondingauthor: true
  qualifications: 29018102
- familyname: Wang
  othernames: Yusen
  address: Monash Universty
  email: "ywa136@student.monash.edu"
  correspondingauthor: true
  qualifications: 27496538
- familyname: Tsakis
  othernames: Alexi
  address: Monash Universty
  email: "atsa0003@student.monash.edu"
  correspondingauthor: true
  qualifications: 29682525
department: Our consultancy \newline Alexi  &\newline Siyi  &\newline Yusen
organization: Australian Government
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    latex_engine: lualatex
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
    in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r , echo = FALSE, include=FALSE, message = FALSE, warning = FALSE}
# Libraries
set.seed(96024)

library(tidyverse)
library(readr)
library(kableExtra)
library(bookdown)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.pos = 'H', echo = FALSE, message = FALSE, warning = FALSE)
```

{Section 1}

####Compulsory Seatbelts 1983

```{r, echo = FALSE, message = FALSE, warning = FALSE}
Seatbelts_data = data.frame(as.matrix(Seatbelts),date=time(Seatbelts))
```

```{r DriversKilled, fig.cap = "Drivers killed in accidents over time", echo = FALSE, message = FALSE, warning = FALSE}
ggplot(data=Seatbelts_data, aes(x=date,y=DriversKilled,col=law))+geom_line()+geom_smooth()
```

```{r FrontKilled, fig.cap = "Frontseat passengers killed in accidents over time", echo = FALSE, message = FALSE, warning = FALSE}
ggplot(data=Seatbelts_data, aes(x=date,y=front,col=law))+geom_line()+geom_smooth()
```

```{r RearKilled, fig.cap = "Backseat passengers killed in accidents over time", echo = FALSE, message = FALSE, warning = FALSE}
ggplot(data=Seatbelts_data, aes(x=date,y=rear,col=law))+geom_line()+geom_smooth()
```

```{r VanKilled, fig.cap = "Light goods vehicle drivers killed in accidents over time", echo = FALSE, message = FALSE, warning = FALSE}
ggplot(data=Seatbelts_data, aes(x=date,y=VanKilled,col=law))+geom_line()+geom_smooth()
```

```{r DriverTable, echo = FALSE, message = FALSE, warning = FALSE}
Seatbelts_data %>%
  group_by(law) %>%
  summarise(DriversKilled_Average=mean(DriversKilled)) 
```

```{r RearTable, echo = FALSE, message = FALSE, warning = FALSE}
Seatbelts_data %>%
  group_by(law) %>%
  summarise(Rear_Average=mean(rear)) 
```

```{r FrontTable, echo = FALSE, message = FALSE, warning = FALSE}
Seatbelts_data %>%
  group_by(law) %>%
  summarise(Front_Average=mean(front)) 
```

```{r VanTable, echo = FALSE, message = FALSE, warning = FALSE}
Seatbelts_data %>%
  group_by(law) %>%
  summarise(Van_Average=mean(VanKilled)) 
```

In this section we look at the effect of seatbelts on driver deaths in the UK. Front seatbelts were compulsory on all new cars registered in the UK from 1968, however it was not required for them to be worn until 1983. In this section, we will answer the question - did mandatory wearing of frontrow seatbelts prevent deaths for drivers and passengers?

First, to answer if any relationships shown are causation and not simply correlation, we ask - why do seatbelts prevent deaths? According to the National Highway Traffic Safety Administration, the primary reasons why seatbelts prevent death is that "Buckling up helps keep you safe and secure inside your vehicle, whereas not buckling up can result in being totally ejected from the vehicle in a crash, which is almost always deadly." (NHTSA 2021)

These scenarios are referred to as "secondary collisions" - i.e. a crash event that results from the impact of an initial collison, such as ejection from a vehicle, or even a high speed impact with objects inside the vehicle.

{Section 2}

{Section 3}
